### 2020-12-07 ch02_scientific_background/scientific_background.tex
Added climate feedback figure and adapted margins

\begin{figure}[t]
  \centering
  \includegraphics[width=\FigureWidth{}]{
    ch02_scientific_background/figs/cmip6.pdf}
  \caption{Schematic of the experiment design of Phase 6 of the \acl{CMIP}
    (\acs{CMIP}6). The center of the circle illustrates the four \acs{DECK}
    (\acl{DECK}) experiments and the \acs{CMIP}6 historical simulation. The
    circular sectors show additional science themes that can be explored
    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}. Taken from
    \textcite{Simpkins2017}.}
  \label{fig:02:cmip6}
\end{figure}

To increase diversity and answer more scientific questions, \acs{CMIP}6 models
can participate in the so-called \acs{CMIP}6-Endorsed \acp{MIP}, of which
\acs{CMIP}6 offers 21 (see circular sectors in \cref{fig:02:cmip6}). Some
\acp{MIP} offer additional experiments to explore specific aspects of the Earth
system, like the \ac{C4MIP} which focuses on the carbon cycle
\autocite{Jones2016} or the \ac{CFMIP} which focuses on the evaluation of cloud
feedbacks \autocite{Webb2017}. Other \acp{MIP} allow the assessment of future
climate change. An example is the \ac{ScenarioMIP}, which provides common
experiments that simulate different possible futures \autocite{ONeill2016}.
These experiments are based on the so-called \acp{SSP}, a set of alternative
pathways of future societal development \autocite{ONeill2013, ONeill2017}. For
each experiment, a set of emissions and land use changes is calculated from the
\acp{SSP} \autocite{Riahi2017} which are then used to force the global climate
models. For \ac{ScenarioMIP}, five different \acp{SSP} are considered, ranging
from \ac{SSP}1 (sustainability) to \ac{SSP}5 (fossil-fuel development). Each
\ac{SSP} is combined with a climate outcome (measured as radiative forcing in
the year 2100) based on a particular forcing pathway that \acp{IAM} have shown
to be feasible. For example, \ac{SSP}5-8.5 represents a scenario based on a
fossil-fuel development with a radiative forcing of $8.5 \unit{W m^{-2}}$ in
2100 while \ac{SSP}1-2.6 represents a sustainable future with a radiative
forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two other main scenarios
(called \emph{Tier 1} experiments in \ac{ScenarioMIP}) are the \ac{SSP}2-4.5
and \ac{SSP}3-7.0 scenarios. In contrast to the \ac{ScenarioMIP} experiments,
the corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012}, the so-called
\acp{RCP}, only used the radiative forcing in 2100 as only dimension to
describe the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5, \ac{RCP}2.6,
\etc{}).

@@

These differences in the climate models also give rise to different intensities
of \emph{climate feedbacks} (or even their presence/absence) in the models. A
climate feedback is a mechanism that either amplifies (\emph{positive
  feedback}) or weakens (\emph{negative feedback}) the effect of an external
forcing. An example of a positive feedback is the water vapor feedback, where
the increased surface temperature (caused by anthropogenic forcing) leads to an
enhanced evaporation of water which enhances the amount of water vapor in the
atmosphere. Since water vapor itself is a powerful \ac{GHG}, this amplifies the
effect of the anthropogenic forcing by further increasing the surface
temperatures \autocite{Cubasch2013}. Further examples and a mathematical
description of feedbacks are given in
\cref{sec:02:feedbacks_and_climate_sensitivity}.

@@

\section{Feedback Processes and Climate Sensitivity}
\label{sec:02:feedbacks_and_climate_sensitivity}

@@

>>>>

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch02_scientific_background/figs/cmip6.pdf}
  \caption{Schematic of the experiment design of Phase 6 of the \acl{CMIP}
    (\acs{CMIP}6). The center of the circle illustrates the four \acs{DECK}
    (\acl{DECK}) experiments and the \acs{CMIP}6 historical simulation. The
    circular sectors show additional science themes that can be explored
    through the 21 \acs{CMIP}6-Endorsed \acfp{MIP}. Taken from
    \textcite{Simpkins2017}.}
  \label{fig:02:cmip6}
\end{figure}

To increase diversity and answer more scientific questions, \acs{CMIP}6 models
can participate in the so-called \acs{CMIP}6-Endorsed \acp{MIP}, of which
\acs{CMIP}6 offers 21 (see circular sectors in \cref{fig:02:cmip6}). Some
\acp{MIP} offer additional experiments to explore specific aspects of the Earth
system, like the \ac{C4MIP} which focuses on the carbon cycle
\autocite{Jones2016} or the \ac{AerChemMIP} which focuses on aerosol chemistry
\autocite{Collins2017}. Other \acp{MIP} allow the assessment of future climate
change. An example is the \ac{ScenarioMIP}, which provides common experiments
that simulate different possible futures \autocite{ONeill2016}. These
experiments are based on the so-called \acp{SSP}, a set of alternative pathways
of future societal development \autocite{ONeill2013, ONeill2017}. For each
experiment, a set of emissions and land use changes is calculated from the
\acp{SSP} \autocite{Riahi2017} which are then used to force the global climate
models. For \ac{ScenarioMIP}, five different \acp{SSP} are considered, ranging
from \ac{SSP}1 (sustainability) to \ac{SSP}5 (fossil-fuel development). Each
\ac{SSP} is combined with a climate outcome (measured as radiative forcing in
the year 2100) based on a particular forcing pathway that \acp{IAM} have shown
to be feasible. For example, \ac{SSP}5-8.5 represents a scenario based on a
fossil-fuel development with a radiative forcing of $8.5 \unit{W m^{-2}}$ in
2100 while \ac{SSP}1-2.6 represents a sustainable future with a radiative
forcing of $2.6 \unit{W m^{-2}}$ in the year 2100. The two other main scenarios
(called \emph{Tier 1} experiments in \ac{ScenarioMIP}) are the \ac{SSP}2-4.5
and \ac{SSP}3-7.0 scenarios. In contrast to the \ac{ScenarioMIP} experiments,
the corresponding \acs{CMIP}5 counterparts \autocite{Taylor2012}, the so-called
\acp{RCP}, only used the radiative forcing in 2100 as only dimension to
describe the possible futures (\eg{} \ac{RCP}8.5, \ac{RCP}4.5, \ac{RCP}2.6,
\etc{}).

@@

These differences in the climate models also give rise to different intensities
of \emph{climate feedbacks} (or even their presence/absence) in the models. A
climate feedback is a mechanism that either amplifies (\emph{positive
  feedback}) or weakens (\emph{negative feedback}) the effect of an external
forcing. An example of a positive feedback is the water vapor feedback, where
the increased surface temperature (caused by anthropogenic forcing) leads to an
enhanced evaporation of water which enhances the amount of water vapor in the
atmosphere. Since water vapor itself is a powerful \ac{GHG}, this amplifies the
effect of the anthropogenic forcing by further increasing the surface
temperatures \autocite{Cubasch2013}. Further examples and a mathematical
description of feedbacks are given in
\cref{subsec:02:climate_feedbacks}.

@@

\section{Climate Sensitivity}
\label{sec:02:climate_sensitivity}

Climate sensitivity refers to the change in the \ac{GSAT} that results from a
change in the radiative forcing. In other words, it describes how sensitive the
climate system is to an external forcing. The source of this forcing might
either be natural (changes in the solar activity, volcanic eruptions, \etc{})
or anthropogenic (emissions of \acp{GHG}, land use changes, \etc{}).

\subsection{Climate Feedbacks}
\label{subsec:02:climate_feedbacks}

As already described in \cref{subsec:02:source_of_uncertainties}, the effects
of an external forcing acting on the climate system can additionally be
amplified or weakened by climate feedbacks. Thus, feedback processes play a
crucial role in the evaluation of climate sensitivity.

\begin{figure}[t]
  \centering
  \includegraphics[width=\LargeFigureWidth{}]{
    ch02_scientific_background/figs/feedbacks_overview.pdf}
  \caption{Climate feedbacks and corresponding time scales. \enquote{+} refers
    to positive feedbacks, which amplify the effect of the external forcing
    (\eg{} the water vapor feedback). \enquote{-} refers to negative feedbacks,
    which weaken the effect of the external forcing (\eg{} the longwave
    radiation feedback). \enquote{+/-} refers to feedbacks which might be
    either positive or negative (\eg{} the cloud feedback). The smaller box
    highlights the large differences in time scales for the various feedbacks.
    Taken from \textcite{Cubasch2013}.}
  \label{fig:02:feedbacks_overview}
\end{figure}

@@

- Timescales of feedbacks

- Longwave radiation feedback = Planck feedback
